Propositions as Sessions, Mechanically

### 2015-04-25 paper/presentation.tex
Commuting conversions more in-depth (too much?).
Added GV example plus some draft bits that I forgot to commit.

@@

@@

\begin{frame}[fragile]
\oldframetitle{Principal Cut Reduction ($\beta_{!?}$)}
\begin{tabular}{c}
\begin{coq}
Lemma reduce_spawn:
  forall P Q A dA $\Gamma$
         (DUA: dual_props A dA)
         (WT: $\nu$ !A.(!0(A).P $\mid$ ?0[dA].Q) $\tpcp$ $\Gamma$),
    $\nu$ A.(P $\mid$ Q) $\tpcp$ $\Gamma$.
\end{coq}
\\~\\
\begin{math}
\begin{array}{c}
\small
\colored
\inference{
  \inference{
    \tm{P} \tpvdash \tp{{?\Gamma} \comma \tmof{y}A}
  }{!}{
    \tm{!x(y).P} \tpvdash \tp{{?\Gamma} \comma \tmof{x}{!A}}
  }
  &
  \inference{
    \tm{Q} \tpvdash \tp{\Delta \comma \tmof{y}\dualbt{A}}
  }{?}{
    \tm{?x[y].Q} \tpvdash \tp{\Delta \comma \tmof{x}{?\dualbt{A}}}
  }
}{}{
  \tm{\nu x \Of{!A}.(!x(y).P \mid {?x[y].Q})} \tpvdash
    \tp{{?\Gamma} \comma \Delta}
}
\becomes
\inference{
  \tm{P} \tpvdash \tp{{?\Gamma} \comma \tmof{y}A}
  \\
  \tm{Q} \tpvdash \tp{\Delta \comma \tmof{y}\dualbt{A}}
}{}{
  \tm{\nu y\Of{A}.(P \mid Q)} \tpvdash \tp{{?\Gamma} \comma \Delta}
}
\normalsize
\end{array}
\end{math}
\end{tabular}
\note{
An example of a principal cut reduction: cut of server accept against client
request. Comparing the formal and informal versions, the informal version
describes the cut reductions using derivation trees while this structure is
left implicit in the Coq encoding.
}
\end{frame}

\begin{frame}
\[\colored
\begin{array}{c}
(\beta_{!?}) \quad
\tm{\nu x \Of{!A}.(!x(y).P \mid {?x[y].Q})}
\quad \becomes \quad
\tm{\nu y\Of{A}.(P \mid Q)}
\\~\\
(\kappa_{\otimes1}) \quad
\tm{\nu z\Of{C}.(x[y].(P \mid Q) \mid R)}
\quad \becomes \quad
\tm{x[y].(\nu z\Of{C}.(P \mid R) \mid Q)}
\end{array}
\]
\end{frame}

\begin{frame}[fragile]
\begin{coq}
Theorem proc_cut_elim:
  forall P $\Gamma$
         (WT: P $\tpcp$ $\Gamma$),
    exists Q, P $\becomes^\star$ Q /\ $\sim$ is_cut Q.
\end{coq}
\note{
mention star being the reflexive and transitive closure of reduction

Well-typed relation alone is insufficient to prove the theorem. The cut typing
rule does not specify the subprocesses are related i.e. are eventually dual
logical operators which would be necessary to proceed using the current
definition.

It seems necessary to define a termination measure on the length of the
derivation but this has not yet been done. Additionally, work on formalising
logics with cut elimination take an entirely different approach favouring an
explicit representation of derivations as trees. Adopting this approach for CP
may prove fruitful not just in this instance but also in other areas of the
development (what? contraction?)

@@

\begin{frame}
GV
\note{
GV has built-in support for session types. consider the rule for sending a
value of type T along a channel N of type !T.S.
}
\end{frame}

@@

\begin{frame}
something about a translation
\note{
the translation to CP is
presented. blah blah something about the session types being a little odd or
overlook this probably
}
\end{frame}

@@

\begin{frame}
problems for cut elimination
\note{
sessions correspond to propositions
processes correspond to proofs
cut elimination corresponds to communication

cut elimination is not currently proven since we require a termination measure
on the length of a derivation and a relationship between the subprocesses of a
cut. other work based on formalising logics where the derivations are explicit
may be inspirational
}
\end{frame}

\begin{frame}
problems of the translation
\note{
the translation seems to require a duplication of the specifications for cp
and gv typing rules not done

what can we show them for the failures?

here are some issues and things learnt! -- our guidelines are all about
modularisation, representation, library reuse, automation!
}
\end{frame}

@@

\begin{frame}
\begin{itemize}
\item difficulties in formalisation
\item implications
\item future work
\end{itemize}
\end{frame}

>>>>

% GV/CP example listing details
\lstdefinelanguage{GV}{
    mathescape=true,
    morekeywords=[1]{case, connect, in, let, of, receive, select, send,
      terminate, to, with},
    sensitive=false,
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]",
    basicstyle=\HappyMed,
    keywordstyle=[1]{\color{magenta}},
}

\lstnewenvironment{gv}{\lstset{language=GV}}{}

@@

\begin{frame}[fragile]
\oldframetitle{GV Example}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} cc}
Process A ($A_x$) & Process B ($B_x$)\\
$\colored\tp{\inpt{Int}{\outpt{String}{\key{end}_!}}}$ ($\colored\tp{S}$) &
$\colored\tp{\outpt{Int}{\inpt{String}{\key{end}_?}}}$
  ($\colored\tp{\overline{S}}$)\\
\begin{gv}
let (m,$x_1$) = receive $x$ in
send thanks $x_1$
\end{gv} &
\begin{gv}
let $x_1$ = send gift $x$ in
let (resp,$x_2$) = receive $x_1$ in
terminate $x_2$
\end{gv}
\end{tabular*}
\[\colored
\inference{
  \Tp{thanks:\tp{String} \comma x:\tp{S}} \Tpvdash A_x:\tp{\key{end}_!} &
  \Tp{gift:\tp{Int} \comma x:\tp{\overline{S}}} \Tpvdash B_x:\tp{\key{Unit}}
}{Connect}{
  \Tp{thanks:\tp{String} \comma gift:\tp{Int}} \Tpvdash
    \Tm{with}~x~\Tm{connect}~A_x~\Tm{to}~B_x : \tp{Unit}
}
\]
\end{frame}

@@

\begin{frame}[fragile]
\oldframetitle{Principal Cut Reduction ($\beta_{!?}$): Formal vs. Informal}
\begin{tabular}{c}
\begin{coq}
Lemma reduce_spawn:
  forall P Q A dA $\Gamma$
         (DUA: dual_props A dA)
         (WT: $\nu$ !A.(!0(A).P $\mid$ ?0[dA].Q) $\tpcp$ $\Gamma$),
    $\nu$ A.(P $\mid$ Q) $\tpcp$ $\Gamma$.
\end{coq}
\\~\\
\begin{math}
\begin{array}{c}
\small
\colored
\inference{
  \inference{
    \tm{P} \tpvdash \tp{{?\Gamma} \comma \tmof{y}A}
  }{!}{
    \tm{!x(y).P} \tpvdash \tp{{?\Gamma} \comma \tmof{x}{!A}}
  }
  &
  \inference{
    \tm{Q} \tpvdash \tp{\Delta \comma \tmof{y}\dualbt{A}}
  }{?}{
    \tm{?x[y].Q} \tpvdash \tp{\Delta \comma \tmof{x}{?\dualbt{A}}}
  }
}{}{
  \tm{\nu x \Of{!A}.(!x(y).P \mid {?x[y].Q})} \tpvdash
    \tp{{?\Gamma} \comma \Delta}
}
\becomes
\inference{
  \tm{P} \tpvdash \tp{{?\Gamma} \comma \tmof{y}A}
  \\
  \tm{Q} \tpvdash \tp{\Delta \comma \tmof{y}\dualbt{A}}
}{}{
  \tm{\nu y\Of{A}.(P \mid Q)} \tpvdash \tp{{?\Gamma} \comma \Delta}
}
\normalsize
\end{array}
\end{math}
\end{tabular}
\note{
An example of a principal cut reduction: cut of server accept against client
request. Comparing the formal and informal versions, the informal version
describes the cut reductions using derivation trees while this structure is
left implicit in the Coq encoding.
}
\end{frame}

\begin{frame}[fragile]
\oldframetitle{Commuting Conversion ($\kappa_{\otimes1}$): Formal vs.
  Informal}
\begin{tabular}{c}
\begin{coq}
Lemma reduce_cc_multi_one:
  forall P Q R (x:atom) A B $\Gamma$ (LCQ: lc_proc Q)
         (WT: $\nu$ A.(x[B].(P $\mid$ Q) $\mid$ R) $\tpcp$ $\Gamma$),
    x[B].($\nu$ A.($\permb{0}{1}{P}$ $\mid$ R) $\mid$ Q) $\tpcp$ $\Gamma$.
\end{coq}
\\~\\
\begin{math}
\colored
\begin{array}{c}

% \spacer
\vcenter{
  \inference{
    \inference{
      \tm{P} \tpvdash \tp{\Gamma \comma \tmof{y}A \comma \tmof{z}C}
      &
      \tm{Q} \tpvdash \tp{\Delta \comma \tmof{x}B}
    }{\otimes}{
      \tm{x[y].(P \mid Q)} \tpvdash
        \tp{\Gamma \comma \Delta \comma \tmof{x}A \otimes B \comma \tmof{z}C}
    }
    &
    \tm{R} \tpvdash \tp{\Theta \comma \tmof{z}C^\bot}
  }{Cut}{
    \tm{\nu z\Of{C}.(x[y].(P \mid Q) \mid R)} \tpvdash
      \tp{\Gamma \comma \Delta \comma \Theta \comma \tmof{x}A \otimes B}
  }
}
\quad \becomes \hfill \\~\\
\hfill
\vcenter{
  \inference{
    \inference{
      \tm{P} \tpvdash \tp{\Gamma \comma \tmof{y}A \comma \tmof{z}C}
      &
      \tm{R} \tpvdash \tp{\Theta \comma \tmof{z}C^\bot}
    }{Cut}{
      \tm{\nu z\Of{C}.(P \mid R)} \tpvdash
        \tp{\Gamma \comma \Theta \comma \tmof{y}A}
    }
    &
    \tm{Q} \tpvdash \tp{\Delta \comma \tmof{x}B}
  }{\otimes}{
    \tm{x[y].(\nu z\Of{C}.(P \mid R) \mid Q)} \tpvdash
      \tp{\Gamma \comma \Delta \comma \Theta \comma \tmof{x}A \otimes B}
  }
}

\end{array}
%% \colored
%% (\kappa_{\otimes1}) \quad
%% \tm{\nu z\Of{C}.(x[y].(P \mid Q) \mid R)}
%% \quad \becomes \quad
%% \tm{x[y].(\nu z\Of{C}.(P \mid R) \mid Q)}
\end{math}
\end{tabular}
\note{

(well, that looks ridiculous.)\\~\\

Comparing formal and informal versions of a commuting conversion for output
reveals some issues. it is necessary to stipulate that Q is locally closed to
ensure that the P subprocess receives the channel from the outer cut as part
of its environment -- since the environment splitting is in general
non-deterministic. Additionally, we need to permute the binders of P in the
conclusion and this is necessary for the majority of the commuting conversion
rules.

\begin{frame}[fragile]
\oldframetitle{Top-Level Cut Elimination}
\begin{coq}
Theorem proc_cut_elim:
  forall P $\Gamma$
         (WT: P $\tpcp$ $\Gamma$),
    exists Q, P $\becomes^\star$ Q /\ $\sim$ is_cut Q.
\end{coq}
\note{
mention star being the reflexive and transitive closure of reduction

Well-typed relation alone is insufficient to prove the theorem. The rule for
cut does not specify the subprocesses are related i.e. are eventually dual
logical operators which would be necessary to proceed using the current
definition.

It seems necessary to define a termination measure on the length of the
derivation but this has not yet been done. Additionally, work on formalising
logics with cut elimination take an entirely different approach favouring an
explicit representation of derivations as trees. Adopting this approach for CP
may prove fruitful not just in this instance but also in other areas of the
development such as specifying environment splitting as in the rules for
commuting conversions.

@@

\begin{frame}
GV
\note{

Typing rules again? Very similar to CP.\\~\\
Representation issues?\\~\\
GV has built-in support for session types. consider the rule for sending a
value of type T along a channel N of type !T.S.

@@

\begin{frame}
something about a translation

\note{

the translation to CP is
presented. blah blah something about the session types being a little odd or
overlook this probably

the translation seems to require a duplication of the specifications for cp
and gv typing rules not done

@@

\begin{frame}
\frametitle{Issues and Guidelines}
\begin{itemize}
\item representation
\item modularisation
\item automation
\item reuse
\end{itemize}
\note{
  representation, modularisation(?), automation, reuse
}
\end{frame}

@@

\begin{frame}
\oldframetitle{Future Work}
\begin{itemize}
\item difficulties in formalisation
\item implications
\item future work
\end{itemize}
\note{
